---
title: "Soil Mapping"
output:
  html_document:
    css: html-md-01.css
    fig_caption: yes
    highlight: haddock
    number_sections: yes
    theme: spacelab
    toc: yes
    toc_float: FALSE
    collapsed: no
---

```{r setup, include=FALSE}
options(width = 105)
knitr::opts_chunk$set(dev='png', dpi=300,cache = TRUE)
pdf.options(useDingbats = TRUE)
```

# Introduction #
This RMarkdown document describes reading and plotting a netCDF file.

## Load packages ##
```{r load_packages, message=FALSE}
# load packages
library(ggplot2)
library(ncdf4)
library(raster)
library(rasterVis)
library(RColorBrewer)
library(ggthemes)
library(ncdf.tools)
library(zoo)
library(sf)
```
# Read Data #
## Shapefile ##
### Read a Shapefile ###
Read world countries shapefile
```{r read_shp}
shp_path <- "C:\\Users\\jonjo\\Documents\\projects\\geog490\\data\\"
shp_name <- "ne_110m_admin_0_countries.shp"
shp_file <- paste(shp_path,shp_name,sep="")

world_shp <- read_sf(shp_file)
world_outline <- as(st_geometry(world_shp), Class = "Spatial")
```
### Plot the shapefile ###
```{r plot_shapefile}
# plot the world outlines
plot(world_outline,col="gray80",lwd=1)
```

## Read Soil ##
Now read in soil moisture data with ncdf4 package
```{r read_data}
nc_path <- "C:\\Users\\jonjo\\Documents\\geog490\\data\\"
nc_name <- "soilw.mon.ltm.v2.nc"
nc_file <- paste(nc_path,nc_name,sep="")

soil_raster <- raster(nc_file,varname="soilw")

soil <- rotate(soil_raster) # Convert the coordinates of raster to -180-180 degrees
soil
```

# Plot the data #
Finally, plot alpha using the `levelplot()` function from `rasterVis`:
```{r plot_data}
# rasterVis plot
maptheme <- rasterTheme(region=brewer.pal(8,"YlOrBr"))
plt <- levelplot(soil, margin=F, par.settings=maptheme)
plt + latticeExtra::layer(sp.lines(world_outline, col="gray30", lwd=0.5))
```

# Plot the long term mean for each month #
```{r plot_months}
NC_soil_ltm <- stack(nc_file,varname="soilw") # Raster stck object
NC_soil_ltm <- rotate(NC_soil_ltm) # Rotate the raster object coordinates
names(NC_soil_ltm) <- c("Dec","Jan","Feb","March","April","May","June","July","August","September","October","November")
#pngfile <- "NC_soil_ltm.png"
#png(pngfile, width=729, height=729) # open the file
cutpts <- c(0,100,200,300,400,500,600,700) # Set the cut points at the range of values of the dataset
cutpts
plt <- levelplot(NC_soil_ltm , at=cutpts, cuts=8, pretty=T, 
                 col.regions=(brewer.pal(8,"YlOrBr")))
plt + latticeExtra::layer(sp.lines(world_outline, col="black", lwd=0.5))
```
